treats manual

### 2024-08-28 inst/gitbook/_book/treats_manual.tex
book up

  pdfauthor={Thomas Guillerme (guillert@tcd.ie)},

@@

\title{treats manual}
\author{Thomas Guillerme (\href{mailto:guillert@tcd.ie}{\nolinkurl{guillert@tcd.ie}})}
\date{2024-05-15}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Simulate disparity and diversity}
\NormalTok{my\_data \textless{}{-}}\StringTok{ }\KeywordTok{treats}\NormalTok{(}\DataTypeTok{bd.params =}\NormalTok{ my\_params,}
                \DataTypeTok{stop.rule =}\NormalTok{ my\_stop\_rule,}
                \DataTypeTok{traits    =}\NormalTok{ my\_trait)}
\end{Highlighting}
\end{Shaded}

@@

They use a dataset from \citet{beckancient2014} that is a discrete morphological space containing the ordination of the around 400 morphological characters for 50 species and estimated for their 49 descendants into an ordinated shapespace of 99 elements and 97 dimensions.

@@

\begin{verbatim}
## $main
## $main$A
## $main$A$process
## function (x0 = 0, edge.length = 1, Sigma = diag(length(x0)), 
##     ...) 
## {
##     return(t(MASS::mvrnorm(n = 1, mu = x0, Sigma = Sigma * edge.length, 
##         ...)))
## }
## <bytecode: 0x609b29299e58>
## <environment: namespace:treats>
## 
## $main$A$start
## [1] 0
## 
## $main$A$trait_id
## [1] 1
## 
## 
## 
## $background
## NULL
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 1 trait for 1 process (A) with one starting value (0).
## process A uses the following extra argument: fun;
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 1 trait for 1 process (A) with one starting value (0).
## process A uses the following extra arguments: fun,min,max;
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 2 traits for 2 processes (A, B) with one starting value (0).
## process A uses the following extra argument: fun;
## process B uses the following extra arguments: fun,min,max;
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 3 traits for 3 processes (A, B, C) with one starting value (0).
## process B uses the following extra argument: fun;
## process C uses the following extra arguments: fun,min,max;
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 3 traits for 3 processes (Normal, LogNormal, Uniform(1,100)) with one starting value (0).
## process LogNormal uses the following extra argument: fun;
## process Uniform(1,100) uses the following extra arguments: fun,min,max;
\end{verbatim}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-56-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-57-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-58-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-59-1.pdf}

@@

@@

@@

\hypertarget{mapping-traits-on-a-tree-map.traits}{%
\section{\texorpdfstring{Mapping traits on a tree: \texttt{map.traits}}{Mapping traits on a tree: map.traits}}\label{mapping-traits-on-a-tree-map.traits}}

@@

@@

\begin{verbatim}
## $waiting
## $waiting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(rexp(1, sum(lineage$n * (bd.params$speciation + bd.params$extinction))))
## }
## <bytecode: 0x609b211b6680>
## <environment: namespace:treats>
## 
## $waiting$internal
## NULL
## 
## 
## $selecting
## $selecting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(sample(lineage$n, 1))
## }
## <bytecode: 0x609b211694e0>
## <environment: namespace:treats>
## 
## $selecting$internal
## NULL
## 
## 
## $speciating
## $speciating$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(runif(1) <= (bd.params$speciation/(bd.params$speciation + 
##         bd.params$extinction)))
## }
## <bytecode: 0x609b21164168>
## <environment: namespace:treats>
## 
## $speciating$internal
## NULL
## 
## 
## $call
## $call$waiting
## $call$waiting$fun
## [1] "default"
## 
## 
## $call$selecting
## $call$selecting$fun
## [1] "default"
## 
## 
## $call$speciating
## $call$speciating$fun
## [1] "default"
\end{verbatim}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-73-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-76-1.pdf}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-79-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-82-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-90-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-91-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-92-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-94-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Making a list of three traits}
\NormalTok{list\_of\_traits \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{process =} \KeywordTok{c}\NormalTok{(no.process, BM.process, OU.process), }\DataTypeTok{trait.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"No process (normal)"}\NormalTok{, }\StringTok{"Brownian motion"}\NormalTok{, }\StringTok{"Ornstein{-}Uhlenbeck"}\NormalTok{))}
\CommentTok{\#\# Plotting each trait separately}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{1}\NormalTok{)}
\CommentTok{\#\# Using different colours options}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{2}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"purple"}\NormalTok{, }\StringTok{"pink"}\NormalTok{))}
\CommentTok{\#\# Not using the default plot name}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{3}\NormalTok{, }\DataTypeTok{main =} \StringTok{"OU process"}\NormalTok{,}
     \DataTypeTok{cent.tend =}\NormalTok{ sd, }\DataTypeTok{quantiles =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{))}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-102-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-103-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-105-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-106-1.pdf}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Specifying a 3D trait process}
\NormalTok{my\_3D\_trait \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{n =} \DecValTok{3}\NormalTok{, )}
\CommentTok{\#\# Simulating a birth{-}death tree with that 3D trait}
\NormalTok{my\_data \textless{}{-}}\StringTok{ }\KeywordTok{treats}\NormalTok{(}\DataTypeTok{bd.params =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{extinction =} \FloatTok{0.2}\NormalTok{),}
                \DataTypeTok{stop.rule =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{max.living =} \DecValTok{50}\NormalTok{),}
                \DataTypeTok{traits    =}\NormalTok{ my\_3D\_trait)}
\end{Highlighting}
\end{Shaded}

@@

@@

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpcKuqE6/file37eb79e5b154}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpcKuqE6/file37eb42b1cdb9}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-116-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-120-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-126-1.pdf}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-134-1.pdf}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpcKuqE6/file37eb202f55d4}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-140-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-144-1.pdf}

@@

@@

  \bibliography{../references.bib}

>>>>

  pdfauthor={Thomas Guillerme (t.guillerme@sheffield.ac.uk)},

@@

\title{treats manual}
\author{Thomas Guillerme (\href{mailto:t.guillerme@sheffield.ac.uk}{\nolinkurl{t.guillerme@sheffield.ac.uk}})}
\date{2024-08-28}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Simulate disparity and diversity}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{43}\NormalTok{)}
\NormalTok{my\_data \textless{}{-}}\StringTok{ }\KeywordTok{treats}\NormalTok{(}\DataTypeTok{bd.params =}\NormalTok{ my\_params,}
                  \DataTypeTok{stop.rule =}\NormalTok{ my\_stop\_rule,}
                  \DataTypeTok{traits    =}\NormalTok{ my\_trait)}
\end{Highlighting}
\end{Shaded}

@@

They use a dataset from \citet{beck2014ancient} that is a discrete morphological space containing the ordination of the around 400 morphological characters for 50 species and estimated for their 49 descendants into an ordinated shapespace of 99 elements and 97 dimensions.

@@

\begin{verbatim}
## $main
## $main$A
## $main$A$process
## $main$A$process[[1]]
## function (x0 = 0, edge.length = 1, Sigma = diag(length(x0)), 
##     ...) 
## {
##     return(t(MASS::mvrnorm(n = 1, mu = x0, Sigma = Sigma * edge.length, 
##         ...)))
## }
## <bytecode: 0x56f306f1ce80>
## <environment: namespace:treats>
## 
## 
## $main$A$start
## [1] 0
## 
## $main$A$trait_id
## [1] 1
## 
## 
## 
## $background
## NULL
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 1 trait for 1 process (A) with one starting value (0).
## process A uses the following extra argument: fun.
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 1 trait for 1 process (A) with one starting value (0).
## process A uses the following extra argument: c("fun", "min", "max").
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 2 traits for 2 processes (A, B) with one starting value (0).
## process A uses the following extra argument: fun.
## process B uses the following extra argument: c("fun", "min", "max").
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 3 traits for 3 processes (A, B, C) with one starting value (0).
## process B uses the following extra argument: fun.
## process C uses the following extra argument: c("fun", "min", "max").
\end{verbatim}

@@

\begin{verbatim}
##  ---- treats traits object ---- 
## 3 traits for 3 processes (Normal, LogNormal, Uniform(1,100)) with one starting value (0).
## process LogNormal uses the following extra argument: fun.
## process Uniform(1,100) uses the following extra argument: c("fun", "min", "max").
\end{verbatim}

@@

\hypertarget{linking-traits-together-with-link.traits}{%
\section{\texorpdfstring{Linking traits together with \texttt{link.traits}}{Linking traits together with link.traits}}\label{linking-traits-together-with-link.traits}}

In the \texttt{treats} algorithm, traits are simulated in parallel for improving performances.
Most of the cases this is a (really) good thing!
However, in some cases you might need some of your traits to be simulated sequentially.
For example if you want a specific trait process or options to be dependent on the result of another trait.
You can set up these linked traits using the \texttt{link.traits} function.
This function takes \texttt{base.trait}, a \texttt{"treats"} \texttt{"traits"} object created using \texttt{make.traits} that will be the first trait simulated and the \texttt{next.trait}, a list of \texttt{"treats"} \texttt{"traits"} objects that will be simulated after the \texttt{base.trait} depending on the link type (\texttt{link.type}) and some other arguments (\texttt{link.args}).
So far the \texttt{link.traits} function only accommodates \texttt{link.type\ =\ "conditional"} for conditional traits.

\hypertarget{conditional-traits}{%
\subsection{Conditional traits}\label{conditional-traits}}

Say for example that you have a primary trait (\texttt{base.trait}) that is a discrete binary trait that generates values \texttt{0} or \texttt{1}.
This can correspond to some abiotic trait for example (e.g.~``is the lineage on an island or on the mainland'').
You can first generate this trait using the \texttt{make.traits} function as detailed above (note here we're using a transition matrix as a specific argument that makes the changes from state 0 to state 1 relatively rare and the reversal even more rare).

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Setting up a discrete trait}
\NormalTok{discrete\_trait \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(discrete.process,}
       \DataTypeTok{process.args =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{transitions =} \KeywordTok{matrix}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{3}\NormalTok{, }\FloatTok{0.2}\NormalTok{, }\FloatTok{0.05}\NormalTok{, }\DecValTok{3}\NormalTok{), }\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{)),}
       \DataTypeTok{trait.names  =} \StringTok{"discrete"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Following this trait value, we can create another trait that is conditional to the outcome of this first discrete trait simulation.
In other words we can specify a choice between two processes depending on the result of the \texttt{base.trait}.
For example, we can say that if the lineage evolves on an island (state \texttt{"1"}) it will evolve using a Brownian motion process but if it evolves on the mainland (state \texttt{"0"}), it will evolve using an OU process.

We'll need to first generate these two traits independently using \texttt{make.traits} (using different options):

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# A standard Brownian Motion trait}
\NormalTok{BM\_trait \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{process =}\NormalTok{ BM.process)}
\CommentTok{\#\# An OU trait with a strong alpha}
\NormalTok{OU\_trait \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{process =}\NormalTok{ OU.process, }\DataTypeTok{process.args =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{alpha =} \DecValTok{10}\NormalTok{))}
\CommentTok{\#\# Combining both traits into a list}
\NormalTok{continuous\_trait \textless{}{-}}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"mainland\_case"}\NormalTok{ =}\StringTok{ }\NormalTok{BM\_trait,}
                         \StringTok{"island\_case"}\NormalTok{   =}\StringTok{ }\NormalTok{OU\_trait)}
\end{Highlighting}
\end{Shaded}

We then need to create a list of functions to interpret the results of the first trait (\texttt{discrete\_trait}) so that if the results of \texttt{discrete\_trait} is \texttt{0}, \texttt{BM\_trait} is used, if it's \texttt{1} \texttt{OU\_trait} is used.
This is done by creating a list of functions that evaluate \texttt{x1}, the outcome of \texttt{discrete\_trait}, and return a logical (\texttt{TRUE} or \texttt{FALSE}).
For example \texttt{function(x1)\ \{x1\ !=\ 0\}} will return \texttt{TRUE} if the \texttt{discrete\_trait} is not state \texttt{0}:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Creating a list of functions to choose between either case for the continuous trait}
\NormalTok{select\_continuous \textless{}{-}}\StringTok{ }\KeywordTok{list}\NormalTok{(}\StringTok{"is.mainland"}\NormalTok{ =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x1) \{x1 }\OperatorTok{==}\StringTok{ }\DecValTok{0}\NormalTok{\},}
                          \StringTok{"is.island"}\NormalTok{   =}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x1) \{x1 }\OperatorTok{==}\StringTok{ }\DecValTok{1}\NormalTok{\})}
\end{Highlighting}
\end{Shaded}

Note that this list must be in the same order as the \texttt{continuous\_trait}: the functions will be evaluated sequentially (i.e.~test the function \texttt{is.mainland}, then \texttt{is.island}) and the first one to return \texttt{TRUE} will be used to select the function in \texttt{continuous\_trait}.
For example in our case if the score of \texttt{discrete\_trait} is \texttt{1}, the second function from \texttt{select\_continuous} is \texttt{TRUE} so the second element from \texttt{continuous\_trait} is used (the \texttt{OU\_trait}).

We can then combine all that using \texttt{link.traits}:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Link these traits}
\NormalTok{conditional\_trait \textless{}{-}}\StringTok{ }\KeywordTok{link.traits}\NormalTok{(}\CommentTok{\# The first trait to be evaluated}
                                 \DataTypeTok{base.trait =}\NormalTok{ discrete\_trait,}
                                 \CommentTok{\# The other traits to be evaluated next}
                                 \CommentTok{\# (here a list of two traits)}
                                 \DataTypeTok{next.trait =}\NormalTok{ continuous\_trait,}
                                 \DataTypeTok{link.type  =} \StringTok{"conditional"}\NormalTok{,}
                                 \CommentTok{\# The functions to choose between either}
                                 \CommentTok{\# continuous traits }
                                 \DataTypeTok{link.args  =}\NormalTok{ select\_continuous)}
\end{Highlighting}
\end{Shaded}

And run the usual simulation pipeline with \texttt{treats}:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Running a tree with this conditional trait}
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}
\NormalTok{my\_conditional\_tree \textless{}{-}}\StringTok{ }\KeywordTok{treats}\NormalTok{(}\DataTypeTok{stop.rule =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{max.taxa =} \DecValTok{200}\NormalTok{),}
                              \DataTypeTok{traits    =}\NormalTok{ conditional\_trait)}

\NormalTok{op \textless{}{-}}\StringTok{ }\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{))}
\KeywordTok{plot}\NormalTok{(my\_conditional\_tree, }\DataTypeTok{trait =} \DecValTok{1}\NormalTok{, }\DataTypeTok{main =} \StringTok{"The conditional trait"}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(my\_conditional\_tree, }\DataTypeTok{trait =} \DecValTok{2}\NormalTok{, }\DataTypeTok{main =} \StringTok{"The conditioned trait"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{par}\NormalTok{(op)}
\CommentTok{\#\# The clumpy bit around the value 0 corresponds to the OU trait}
\end{Highlighting}
\end{Shaded}

Of course, you can combine linked traits with other traits like any other traits in \texttt{treats} using \texttt{make.traits} (for example: \texttt{make.traits(process\ =\ my\_new\_trait,\ add\ =\ conditional\_trait)} to generate more complex traits.
In this case the traits are generated in parallel (for speed) apart for the specified linked ones.

@@

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-63-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-64-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-65-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-66-1.pdf}

@@

\hypertarget{map.traits}{%
\section{\texorpdfstring{Mapping traits on a tree: \texttt{map.traits}}{Mapping traits on a tree: map.traits}}\label{map.traits}}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-67-1.pdf}

@@

\begin{verbatim}
## $waiting
## $waiting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(rexp(1, sum(lineage$n * (bd.params$speciation + bd.params$extinction))))
## }
## <bytecode: 0x56f30a0bdef0>
## <environment: namespace:treats>
## 
## $waiting$internal
## NULL
## 
## 
## $selecting
## $selecting$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(sample(lineage$n, 1))
## }
## <bytecode: 0x56f30a0c0e78>
## <environment: namespace:treats>
## 
## $selecting$internal
## NULL
## 
## 
## $speciating
## $speciating$fun
## function (bd.params, lineage = NULL, trait.values = NULL, modify.fun = NULL) 
## {
##     return(runif(1) <= (bd.params$speciation/(bd.params$speciation + 
##         bd.params$extinction)))
## }
## <bytecode: 0x56f30a0c0388>
## <environment: namespace:treats>
## 
## $speciating$internal
## NULL
## 
## 
## $call
## $call$waiting
## $call$waiting$fun
## [1] "default"
## 
## 
## $call$selecting
## $call$selecting$fun
## [1] "default"
## 
## 
## $call$speciating
## $call$speciating$fun
## [1] "default"
\end{verbatim}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-81-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-83-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-84-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-87-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-95-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-96-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-97-1.pdf}

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Making a list of three traits}
\NormalTok{list\_of\_traits \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{process =} \KeywordTok{c}\NormalTok{(no.process, BM.process, OU.process),}
                              \DataTypeTok{trait.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"No process (normal)"}\NormalTok{, }\StringTok{"Brownian motion"}\NormalTok{, }\StringTok{"Ornstein{-}Uhlenbeck"}\NormalTok{))}
\CommentTok{\#\# Plotting each trait separately}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{2}\NormalTok{, }\DecValTok{2}\NormalTok{))}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{1}\NormalTok{)}
\CommentTok{\#\# Using different colours options}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{2}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{"red"}\NormalTok{, }\StringTok{"purple"}\NormalTok{, }\StringTok{"pink"}\NormalTok{))}
\CommentTok{\#\# Not using the default plot name}
\KeywordTok{plot}\NormalTok{(list\_of\_traits, }\DataTypeTok{trait =} \DecValTok{3}\NormalTok{, }\DataTypeTok{main =} \StringTok{"OU process"}\NormalTok{,}
     \DataTypeTok{cent.tend =}\NormalTok{ sd, }\DataTypeTok{quantiles =} \KeywordTok{c}\NormalTok{(}\DecValTok{10}\NormalTok{, }\DecValTok{30}\NormalTok{))}
\KeywordTok{par}\NormalTok{(}\DataTypeTok{mfrow =} \KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{, }\DecValTok{1}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-108-1.pdf}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-111-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-112-1.pdf}

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Specifying a 3D trait process}
\NormalTok{my\_3D\_trait \textless{}{-}}\StringTok{ }\KeywordTok{make.traits}\NormalTok{(}\DataTypeTok{n =} \DecValTok{3}\NormalTok{)}
\CommentTok{\#\# Simulating a birth{-}death tree with that 3D trait}
\NormalTok{my\_data \textless{}{-}}\StringTok{ }\KeywordTok{treats}\NormalTok{(}\DataTypeTok{bd.params  =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{extinction =} \FloatTok{0.2}\NormalTok{),}
                  \DataTypeTok{stop.rule  =} \KeywordTok{list}\NormalTok{(}\DataTypeTok{max.living =} \DecValTok{50}\NormalTok{),}
                  \DataTypeTok{traits     =}\NormalTok{ my\_3D\_trait,}
                  \DataTypeTok{null.error =} \DecValTok{100}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
## Building the tree:..Done.
\end{verbatim}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-114-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-115-1.pdf}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpNKmbbM/file9d7231f94b74}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpNKmbbM/file9d724094c3af}

\hypertarget{using-plot.treats-to-plot-non-treats-objects}{%
\section{\texorpdfstring{Using \texttt{plot.treats} to plot non-treats objects}{Using plot.treats to plot non-treats objects}}\label{using-plot.treats-to-plot-non-treats-objects}}

If you like the plotting style of \texttt{treats} trees and data, you can always use the \texttt{make.treats} function to plot a tree and some traits as a \texttt{"treats"} object.
If you have a dataset that contains trait data for tips and nodes you can transform it into a \texttt{"treats"} object and plot it using the generic \texttt{plot.treats} function:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{set.seed}\NormalTok{(}\DecValTok{1}\NormalTok{)}
\CommentTok{\#\# Generating some random tree}
\NormalTok{my\_tree \textless{}{-}}\StringTok{ }\KeywordTok{makeNodeLabel}\NormalTok{(}\KeywordTok{rcoal}\NormalTok{(}\DecValTok{20}\NormalTok{))}

\CommentTok{\#\# Generating some random discrete data}
\NormalTok{discrete\_data \textless{}{-}}\StringTok{ }\KeywordTok{sample}\NormalTok{(}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{2}\NormalTok{), }\DecValTok{39}\NormalTok{, }\DataTypeTok{replace =} \OtherTok{TRUE}\NormalTok{)}
\KeywordTok{names}\NormalTok{(discrete\_data) \textless{}{-}}\StringTok{ }\KeywordTok{c}\NormalTok{(my\_tree}\OperatorTok{$}\NormalTok{tip.label, my\_tree}\OperatorTok{$}\NormalTok{node.label)}

\CommentTok{\#\# Creating a treats object with this data}
\NormalTok{discrete\_treats \textless{}{-}}\StringTok{ }\KeywordTok{make.treats}\NormalTok{(}\DataTypeTok{tree =}\NormalTok{ my\_tree, }\DataTypeTok{data =}\NormalTok{ discrete\_data)}

\CommentTok{\#\# Plot the data together}
\KeywordTok{plot}\NormalTok{(discrete\_treats, }\DataTypeTok{legend =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-118-1.pdf}

The same goes for continuous data:

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\#\# Generating some random continuous data}
\NormalTok{continuous\_data \textless{}{-}}\StringTok{ }\KeywordTok{rnorm}\NormalTok{(}\DecValTok{39}\NormalTok{)}
\KeywordTok{names}\NormalTok{(continuous\_data) \textless{}{-}}\StringTok{ }\KeywordTok{c}\NormalTok{(my\_tree}\OperatorTok{$}\NormalTok{tip.label, my\_tree}\OperatorTok{$}\NormalTok{node.label)}

\CommentTok{\#\# Creating a treats object with this data}
\NormalTok{continuous\_treats \textless{}{-}}\StringTok{ }\KeywordTok{make.treats}\NormalTok{(}\DataTypeTok{tree =}\NormalTok{ my\_tree, }\DataTypeTok{data =}\NormalTok{ continuous\_data)}

\CommentTok{\#\# Plot the data together}
\KeywordTok{plot}\NormalTok{(continuous\_treats)}
\end{Highlighting}
\end{Shaded}

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-119-1.pdf}

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-127-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-133-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-137-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-141-1.pdf}

@@

\includegraphics[width=6.5in]{../../../../../../tmp/RtmpNKmbbM/file9d727fb11a9c}

@@

@@

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-154-1.pdf}

@@

\includegraphics{treats_manual_files/figure-latex/unnamed-chunk-158-1.pdf}

  \bibliography{../References.bib}
